---
title: "Project 2"
author: "Cameron Farrugia"
date: "4/21/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require("COVID19")
require(dplyr)
require(ggplot2)
require(tidyverse)
```


Making state a Factor because I want to group by that variable
```{r}
covid=covid19("ITA", level = 2)
# write.csv(covid, file="C:/Users/camer/Desktop/DAT512 Stat Big Data/covidOG.csv")
str(covid)
covid$state=factor(covid$state)

```

Since the data is cumulative I want to eliminate all the entries that arn't the most recent data.
```{r}
# Filter to show the most recent data since it is cumlative. 
covid2=covid
covid2=covid2 %>% 
  filter(date == "2020-04-20")
```

Adding in some useful columns and making one a categorical variable
```{r}
covid2=covid2 %>% group_by(state) %>% mutate(recover_rate=recovered/confirmed, Population_category=pop_density)
covid2$Population_category= cut(covid2$Population_category, breaks=c(0,150,300,1000), labels=c("low", "medium", "high"))


```

Shwoing recovery rate based on population area. More factors go into why it shows low density having the higher recovery rate, but I thought it was an interesting point to be made.
```{r}
covid3=covid2 %>% group_by(Population_category) %>% mutate(recover_rate=mean(recover_rate)) %>% distinct(recover_rate)
View(covid3)
ggplot(covid3, aes(x=Population_category, y= recover_rate, fill=Population_category))+geom_col()

```

This shows the percent of overall population in an area that have COVID-19
```{r}
covid4= covid2 %>% mutate(percent_infect= confirmed/pop)
View(covid4)
ggplot(covid4, aes(x=state, y=percent_infect, fill=state, key=FALSE))+geom_col()

```

First attempted linear model. y= confirmed x1=population, x2=population density, x3=Population Category, x4=tests I included these parameters because they seemed the most likely to affect the number of confirmed cases. 
```{r}
model<- lm(confirmed ~ pop+pop_density+Population_category+tests, data=covid4)
summary(model)
```

This model isn't very good because the F statistic is pretty low. This maybe because of the dataset size or the predictors just arn't very good at predicting a confirmed case.